---
title: "covid 19"
author: "Cynthia Tellez"
date: "10/4/2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 
### This is contained in a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Question 1 - Is dying independent of sex




Table of patients:
```{r}

# cross tabulation, patient state by gender

with(kRdata, table(Sex, result))

```

## We create two bar charts - one based on frequency and the other on percent, use ggplot commands to do this

```{r}
barchartGC(~result + Sex,data=kRdata)
barchartGC(~result + Sex,data=kRdata, type="percent")

```

The first plot shows that the amount of people who died is not that great compared to the amount of people who are still alive. We can tell that by looking at the size of the two plots. The second plot shows that the live females is a greater amount than that of the live males. The live females seem to have outnumbered the live males. The amount of dead males is almost double the size of the amount of dead females.

## Numerical results

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```
In the numerical data provided to us, we are able to see accurate numbers in terms of those females and males that have been found dead and alive. The data table above shows that 98.7% of females were found alive after covid, in comparison to the 97% males found alive after covid. This goes to show that the amount of dead females was 1.4% while the amount of dead males was more than doubled with a percent of 3. 

## Inferential results 

The null hypothesis is
$H_0$: There is no difference in the portion of men dying compared to the portion of women. 

$H_A$: There is a difference in the proportion of men dying compared to the women.

```{r}
chisq.test(table1)
#Next we do tiger stats chi sq test
chisqtestGC(table1)
#Next we will do a fisher test
fisher.test(table1)
```

The chi sq test provides us with the exact amount of individuals whom have either lived or passed. It shows that 1831 were females and 1411 males were found alive. It also shows how many of each had died with 25 being females and 44 being males. 

The fisher test gives us the odds of dying. For males we can compute the exact number by using the equation 44/1411 and for females we can use the equation 25/1831. The odds ratio is 44/1411 divided by 25/1831. The odds ratio for the chance of a male to a female dying is 2.3. Males are 2.3 times more likely to die than females. 

## Conclusions

